Faithful Model Explanations through\\ Energy-Constrained Conformal Counterfactuals\footnote{This is a pre-print.}

### 2023-12-16 paper/partials/appendix.tex
appendix done

The core part of our code base is integrated into a larger ecosystem of \texttt{Julia} packages that we are actively developing and maintaining. To avoid compromising the double-blind review process, we only provide a link to an anonymized repository at this stage: \url{https://anonymous.4open.science/r/ECCCo-1252/README.md}.

@@

In our experiments we always generate multiple counterfactuals for each model and generator. Each time the factual and target class is drawn randomly. For each generator and model we choose $n_f=100$ factuals for all of our synthetic and vision data. For tabular data we choose $n_f=25$ because larger values made grid search computationally prohibitive. For vision data, grid search was computationally prohibitve in any case, so hyperparameters were tuned manually. For all other datasets grid search was performed over different combinations of penalty strengths and optimizer steps sizes (details of which can be found in the code). To select the final hyperparameter setting we used the unfaithfulness metric as our criterion.

>>>>

We have built a small Julia package, \texttt{ECCCo.jl}, that was used to perform the analysis in this paper: https://github.com/pat-alt/ECCCo.jl. The package will be absorbed by \texttt{CounterfactualExplanations.jl} in the future~\citep{altmeyer2023explaining}.

@@

In our experiments, we always generate multiple counterfactuals for each model and generator over multiple runs to account for stochasticity. For each generator, model and dataset, we sample a total of $n_f=100$ factuals for a single benchmark, where each time the factual and target class is drawn randomly. This process is repeated over 50 runs for the synthetic and tabular datasets, which yields a total of around 1.5 million counterfactuals per dataset for the final results. For vision datasets, we relied on 10 runs yielding a total of 300,000 counterfactuals. The final results presented in the paper and this appendix show the sample averages and the standard deviations of samples averages across all runs. For grid search, we used single runs for each parameter specification generating 100 counterfactuals per generator and model for synthetic and tabular datasets. For vision data, we generated 10 counterfactuals for each model, generator and parameter specification.

### 2023-12-16 paper/partials/body.tex
appendix done

Generating multiple samples using SGLD thus yields an empirical distribution $\widehat{\mathbf{X}}_{\theta,\mathbf{y}^+}$ that approximates what the model has learned about the input data. While in the context of EBM, this is usually done during training, we propose to repurpose this approach during inference in order to evaluate the faithfulness of model explanations. The appendix provides additional implementation details for any tasks related to energy-based modelling\footnote{The supplementary appendix can be found here: https://github.com/pat-alt/ECCCo.jl.}.

>>>>

Generating multiple samples using SGLD thus yields an empirical distribution $\widehat{\mathbf{X}}_{\theta,\mathbf{y}^+}$ that approximates what the model has learned about the input data. While in the context of EBM, this is usually done during training, we propose to repurpose this approach during inference in order to evaluate the faithfulness of model explanations. The appendix provides additional implementation details for any tasks related to energy-based modelling.
